### 2016-11-26 dissertation/recha.tex
dissertation v2_part normal complete

\section{Motivation}
\subsection{Limitations of Full Scan Algorithms}
The inspiration of this research mainly comes from a journal paper \cite{KallaPartialScan}. 
In that paper, the author proposed an traversal algorithm using concept ``implicit state enumeration".
Concretely, the algorithm is written as follows:

@@

Since a typical sequential circuit usually contains an functionally independent combinational logic
component, the traversal algorithm analyzes the combinational logic and derive the transition 
function for one-step reachability within current time-frame, and extend the result to complete 
execution paths through unrolling. If each state encoding (i.e. exact values in the selected registers) is
explicitly exposed and counted during the unrolling 
procedure, this unrolling is called {\bf explicit unrolling}. In the SIMPSON algorithm, the states cannot be 
directly read in the execution; instead, they are implicitly represented using a conjuction of several 
Boolean formulas. Such techniques differs from explicit unrolling are called {\bf implicit unrolling}.

However, the SIMPSON algorithm proposed by the author is usually not practical. The conjuctions of 
Boolean formulas are stored as BDDs, which is a canonical and convenient structure. Nevertheless, 
the size of BDD is easy to explode when the formulas become too long and too complicated. In the paper,
the author make a compromise between accuracy and cost, and turn to a partial scan technique rather than 
a full scan. In my research work, the aim is to explore a word-level technique which can make a full scan 
available.

\subsection{Word-level Data Flow on Modern Datapaths}

\subsection{Prerequisites of Word-level Techniques}

@@

Conceptually, the state-space of a FSM is traversed in a breadth-first
manner, as shown in Algorithm \ref{alg:BFS}. % \cite{KallaPartialScan}: 
The algorithm operates on the FSM $\mathcal{M} = (\sum, O, S, S^0,
\Delta, \Lambda)$ underlying a sequential circuit. In such cases, the
transition function $\Delta$ and the initial states are represented
and manipulated using Boolean representations such as BDDs or SAT
solvers. The variables $from, reached, to, new$ represent
characteristic functions of sets of states. Starting from the initial
state $from^i = S^0$, the algorithm computes the states reachable in
1-step from $from^i$ in each iteration. In line 4 of algorithm
\ref{alg:BFS}, the {\it image computation} is 
used to compute the reachable states in every execution step.

@@

Let us describe the application of the algorithm on the  FSM circuit
of Fig. \ref{fig:fsm}. {\it We will first describe its operation at the
Boolean level, and then describe how this algorithm can be implemented
using algebraic geometry at word level.}

@@

In Line 1 of the BFS algorithm, assume that the initial state
is $S_3$ in Fig.\ref{fig:fsm}(b), which is encoded as 
$S_3 = \{11\}$. Using Boolean variables $s_0, s_1$ for the present
states, $from^0 = s_0\cdot s_1$ is represented as a Boolean formula.

\begin{Example}
For the circuit in Fig. \ref{fig:fsm} (b), we have the transition
functions of the machine as:

@@

{\bf FSM Traversal at word-level over $\Fkk$:} 
The state transition graph (STG) shown in Fig.\ref{fig:fsm}(a) uses a
2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2,
S_3\}$. We map these states to elements in $\mathbb{F}_{2^2}$, where
$S_0 = 0, S_1 = 1, S_2 = \alpha, S_3 = \alpha+1$. Here, we take 
$P(X) = X^2+X+1$ as the irreducible polynomial to construct
$\mathbb{F}_4$, and $P(\alpha) = 0$ so that $\alpha^2 + \alpha + 1 =
0$.

{\it Initial state:} In Line 1 of Alg.\ref{alg:BFS}, the initial state is
specified by means of a corresponding polynomial $f = \mathcal{F}(S) =
S - 1 - \alpha$. Notice that if we consider the ideal generated by the
initial state polynomial, $I = \langle f\rangle$, then its variety
$V(I) = 1+\alpha$ corresponds to the state encoding $S_3 = \{11\} =
1+\alpha$, where a polynomial in word-level variable $S$ encodes the
initial state. 
% with only one generator $f$, its variety
% $V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 1+\alpha\}$, which equals to $\{1+\alpha\}$, the only
% valid value $S_3$ can take.

@@

{\bf Set operations:} In Lines 5 and 6 of
Alg. \ref{alg:BFS}, we need  \textbf{union}, 
\textbf{intersection} and \textbf{complement} of varieties over
$\mathbb{F}_{2^k}$, for which we again use algebraic geometry concepts.

@@

In Line 5 of Alg. \ref{alg:BFS}, we need to compute the complement of a
set of states. Assume that $J$ denotes a polynomial ideal whose
variety $V(J)$ denotes a set of states. We require the computation of
another polynomial ideal $J'$, such that $V(J')=\overline{V(J)}$. We
show that this computation can be performed using the concept of {\bf
  ideal quotient}:

@@

\begin{proof}
Let $d = gcd(f, g)$. So, $f = df_1 , g = dg_1$ with $gcd(f_1 , g_1 ) =
1$. Note that $f_1 = \frac{f}{gcd(f,g)}$.

@@

Conversely, let $h \in \langle f_1 \rangle$. Then $h = s \cdot f_1$,
where $s \in \Fkk[x]$. So, $hg = hdg_1 = sf_1 dg_1 = sg_1 f \in 
\langle f \rangle$. Therefore, $h \in \langle f : g\rangle$.
\end{proof}

@@

\begin{proof}
Since $\Fkk[x]$ is a principal ideal domain, $ J = \langle g\rangle$
for some polynomial $ g \in \Fkk[x]$. Let $d = gcd(g, x^{2^k} -
x)$. So, $g = dg_1 , x^{2^k} - x = df_1$, with $gcd(f_1 , g_1 ) =
1$. Then $J_0 : J = \langle f_1 \rangle$ by Lemma \ref{lem:gcd}.

@@

So, $x \notin V(g)$ and this concludes the proof.
\end{proof}

@@

With Corollary \ref{cor:complement}, we are ready to demonstrate the
concept of word-level FSM traversal over $\Fkk$ using algebraic
geometry. The algorithm is given in Alg. \ref{alg:univa}. Note that in
the algorithm, $from^i, to^i, new^i$ are {\it univariate polynomials in variables $S$
or $T$} only, due to the fact that they are the result of a GB
computation with an elimination term order, where the bit-level 
variables are abstracted and quantified away.

%% Assume $I = \langle f\rangle  = \langle T^2 +
%% (1+\alpha)\cdot T+\alpha\rangle $,  we can add vanishing polynomial
%% $T^4 - T$ to this ideal such that its variety  $V(I) = \{a\ |\ a \in
%% \mathbb{F}_{2^2}\ and\ f(a) = 0\} = \{1, \alpha\}$. For complement set
%% of variety $\{1, \alpha\}$, the universal set is the variety of ideal
%% of vanishing polynomial $V(\langle T^4-T\rangle ) =
%% \{0,1,\alpha,1+\alpha\}$, 
%% so $\overline{V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )} = V(\langle T^4-T\rangle ) - V(\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle )$,
%% which equals to $V(\langle T^4-T\rangle :\langle T^2 + (1+\alpha)\cdot T+\alpha\rangle ) = V(\langle T^2+(1+\alpha)\cdot T\rangle )$,
%% the result is $\{0,1+\alpha\}$.

@@

\begin{Example}
\label{ex:SMPO}
We apply Algorithm \ref{alg:univa} to the example shown in
Fig. \ref{fig:fsm} to execute the FSM traversal. Let the initial state
$from^0 = \{00\}$ in $\B^2$ or $0 \in \mathbb F _4$. Polynomially, it is
written as $from^0 = S - 0$. In the first iteration, we compose
an ideal $J$ with

@@

>>>>

\section{Motivation}
The motivation of this part of my research consists of many inspirations and observations.
In this dissertation, we will summarize those major ones into 3 topics. First, we start from 
a sequential circuit partial scan paper to discuss the inspirations of our work as well as some 
difficulties to overcome. Secondly, we state the observation that word-level description is 
increasingly important and common in characterizing the data flow on modern large-size datapaths. 
Last but not least, we give instances that some prerequisites for supporting word-level techniques 
are already available. Thus we cover both the necessity and sufficiency of developing such 
a word-level traversal technique in our work.

\subsection{Limitations of Full Scan Algorithms}
The inspiration of this research mainly comes from a journal paper \cite{KallaPartialScan}. 
In that paper, the author proposed an traversal algorithm using concept ``implicit state enumeration".
Concretely, the algorithm is written as follows:

@@

Since a typical sequential circuit usually contains an functionally independent combinational logic
component, the traversal algorithm analyzes the combinational logic and derive the transition 
function for one-step reachability within current time-frame, and extend the result to complete 
execution paths through unrolling. If each state encoding (i.e. exact values in the selected registers) is
explicitly exposed and counted during the unrolling 
procedure, this unrolling is called {\bf explicit unrolling}. In the SIMPSON algorithm, the states cannot be 
directly read in the execution; instead, they are implicitly represented using a conjunction of several 
Boolean formulas. Such techniques differs from explicit unrolling are called {\bf implicit unrolling}.

However, the SIMPSON algorithm proposed by the author is usually not practical. The conjunctions of 
Boolean formulas are stored as BDDs, which is a canonical and convenient structure. Nevertheless, 
the size of BDD is easy to explode when the formulas become too long and too complicated. In the paper,
the author make a compromise between accuracy and cost, and turn to a partial scan technique rather than 
a full scan. In my research work, the aim is to explore a word-level technique which can make a full scan 
available.

\subsection{Word-level Data Flow on Modern Datapath Designs}
The integrity of modern digital circuit designs is very high. For example, processor A10 designed by Apple
integrates 3.3 billion transistors on a $125~mm^2$ chip \cite{AppleA10}. Such high integrity make the silicon 
implementation of large datapaths possible. In recent decades, leading by top-notch central processing units (CPUs)
and high bandwidth memory (HBM), 64-bit or even larger datapaths frequently appear in digital ICs.
Meanwhile, with the development of electrical design automation (EDA) tools, data flow on those datapaths
are directly described by the designer as word-level specifications. Therefore, it will be straightforward 
and beneficial for users if formal verification tools can work on word-level. Moreover, adopting word-level 
techniques will greatly reduce the state space and make verification executes faster by orders of magnitude.

In order to throw light on the advantages using word-level techniques, we pick a typical 
digital circuit component in modern 64-bit MIPS processor as an example.

\begin{Example}[Verification of sequential multiplication hardware]

% fig:MIPS

Figure \ref{fig:MIPS} depicts a sequential multiplication hardware implementation within a 64-bit 
MIPS. Initially, one multiplicand is preloaded to the lower 64 bits of the product registers. 
Iteratively, the last significant bit (LSB) of current (temporary) product is used as flag to activate the ALU 
to add on the other multiplicand. For each iteration the data in product registers shift right by 1 bit.
Finally when the most significant bit (MSB) of preloaded multiplicand arrives the MSB of product registers,
the registers contains the result -- 128 bits product. The behavior can be concluded as following algorithm:

\vspace{0.2in}

\IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
\Indm
 \KwIn{Multiplicand $A,B$}
 \KwOut{Product $C$}
\Indp

  Preload $B$ into lower 64-bit of Product Register $P$\;
  \Repeat{64 Repetitions}
  {
  	\If{Last Bit of Product Register $LSB(P)==1$}
  	{
		$P = P_{1/2}+B$\;
	}
	Right shift $P$\;
  }
\Return{$C=P$}
\caption {Sequential multiplication hardware in 64-bit MIPS}\label{alg:MIPS}
\end{algorithm}
\DecMargin{1em}

\vspace{0.2in}

% List the detailed bit-level property need to check P==P+B
Traditionally, to verify the functional correctness of this multiplier, satisfiability (SAT) based or BDD based 
model checking is applied on basic function units. For example, as a part of functional verification, 
we would like to check ``$P=P_{1/2}+B$" is correctly executed. Then in a model checker we need to add following 
specifications:

\begin{align*}
& en\_ALU \\
\land & s_0 = a_0 \oplus p_0 \\
\land & s_1 = a_1 \oplus p_1 \oplus (a_0 \land p_0) \\
\land & s_2 = a_2 \oplus p_2 \oplus ((a_1 \oplus p_1) \land a_0 \land p_0 \lor a_0 \land p_0) \\
\land & \cdots \\
\land & s_{63} = a_{63} \oplus p_{63} \oplus (c_{63})
\end{align*}

We can see that when checking a single part of the whole structure, the number of clauses 
needed will increase to $k+1$ when using $k$-bit datapath. Considering the formula representing 
carry-in will become longer and longer, the final conjunction of all clauses will contain 
$O(2^k)$ Boolean operators. If by some means we can write the specification with only 3 variables
\begin{equation}
\label{eqn:wordspec}
S=P_{1/2}+B
\end{equation}
The state space will be greatly reduced.
\end{Example}

On the other hand, the user of conventional model checker will have to write all these clauses.
These clauses contain cross-literals, e.g. $s_2$ may associate with $a_1, p_0$, etc. If the user 
is not familiar with the implementation of this adder, those cross-literals will bring inconveniences
and confusions. However, if word-level techniques allow specification like equation \ref{eqn:wordspec},
the verification tool will be very user-friendly and straightforward.

\subsection{Prerequisites of Word-level Techniques}
When given a bit-level netlist, the prerequisites to use word-level techniques are to convert bit-level 
to word-level first. This conversion is usually completed by abstraction techniques.

An old but universally effective abstraction method is {\bf Lagrange's interpolation}. Instead of using 
it in real algebra, it can also be extended to Galois field. Here we use an example to illustrate the 
conversion in Galois field using Lagrange's interpolation.

\begin{Example}[Lagrange's interpolation]

\begin{figure}[H]
\centering{
\includegraphics[width=4in]{newfig/Lagrange.pdf}
\caption{Gate-level netlist for Lagrange's interpolation example}
\label{fig:Lagrange}}
\end{figure}

Assume we are given gate-level netlist shown in figure \ref{fig:Lagrange}. It can be written as 3 
Boolean equations:
\begin{align*}
z_0 &= a_0 \lor a_1 \lor a_2 \\
z_1 &= a_1 \land \neg a_2 \lor a_0 \land \neg a_1 \land a_2 \\
z_2 &= a_1 \lor \neg a_0 \land a_2
\end{align*}
Define 2 word-level variables $A,Z$ as input and output:
$$A = \{a_2a_1a_0\},~Z = \{z_2z_1z_0\}$$
To convert bit-level to word-level, we need to find mapping $\mathbb{B}^3 \rightarrow \mathbb{B}^3$,
or $\F_{2^3} \rightarrow \F_{2^3}$. The latter one, as mentioned in preliminaries, is a polynomial
function in $\F_{2^3}$.

For each element in $\F_{2^3}$, we write down the truth table as follows:
\begin{table}[H]
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$}
\end{table}
Now our objective is to abstract a function over GF $\F_{2^3}$ about word-level variables, i.e. $Z = \Func(A)$.
Recall the definition of Lagrange's interpolation:
\begin{equation}
\label{eqn:Lagrange}
\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
\end{equation}
The geometry meaning of Lagrange's interpolation in real algebra is: given $N$ dots with coordinates $(x_i,y_i)$,
they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
written in the form of equation \ref{eqn:Lagrange}. In this example, although defined in Galois field instead of 
real number field, the essential concept of Lagrange's interpolation keeps the same. We can get 8 ``dots":
\begin{align*}
Uni-form&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
Dot~1&: (0, 0) \gets (000,000) \\
Dot~2&: (1,1) \gets (001,001) \\
Dot~3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
Dot~4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
Dot~5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
Dot~6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
Dot~7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
Dot~8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
\end{align*}
Substitute 8 $(x_i,y_i)$ pairs in equation \ref{eqn:Lagrange} with these 8 ``dots" in $\F_{2^3}$.
The result is a polynomial function with degree no greater than 7:
\begin{align*}
Z =& \Func(A) \\
=& (\alpha^2+\alpha+1) A^7 + (\alpha^2+1) A^6 + \alpha A^5 + (\alpha+1) A^4 \\
&+ (\alpha^2+\alpha+1)A^3+ (\alpha^2+1)A
\end{align*}

\end{Example}

The Lagrange's interpolation theorem also proves the existence of a word-level abstraction for a bit-level netlist.
In practice, Lagrange's interpolation is too weak on scalability. Our approach uses abstraction based on 
Gr\"obner basis with abstraction term order (ATO), which is briefly introduced in preliminaries.

@@

\subsection{Conventional Traversal Method}
Conceptually, the state-space of a FSM is traversed in a breadth-first
manner, as shown in Algorithm \ref{alg:BFS}. % \cite{KallaPartialScan}: 
The algorithm operates on the FSM $\mathcal{M} = (\sum, O, S, S^0,
\Delta, \Lambda)$ underlying a sequential circuit. In such cases, the
transition function $\Delta$ and the initial states are represented
and manipulated using Boolean representations such as BDDs or SAT
solvers. The variables $from, reached, to, new$ represent
characteristic functions of sets of states. Starting from the initial
state $from^i = S^0$, the algorithm computes the states reachable in
1-step from $from^i$ in each iteration. In line 4 of algorithm
\ref{alg:BFS}, the {\it image computation} is 
used to compute the reachable states in every execution step.

@@

Let us describe the application of the algorithm on the  FSM circuit
of figure \ref{fig:fsm}. {\it We will first describe its operation at the
Boolean level, and then describe how this algorithm can be implemented
using algebraic geometry at word level.}

@@

In Line 1 of the BFS algorithm, assume that the initial state
is $S_3$ in figure \ref{fig:fsm}(b), which is encoded as 
$S_3 = \{11\}$. Using Boolean variables $s_0, s_1$ for the present
states, $from^0 = s_0\cdot s_1$ is represented as a Boolean formula.

\begin{Example}
For the circuit in figure \ref{fig:fsm} (b), we have the transition
functions of the machine as:

@@

\subsection{FSM Traversal at word-level over $\Fkk$} 
The state transition graph (STG) shown in figure \ref{fig:fsm}(a) uses a
2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2,
S_3\}$. We map these states to elements in $\mathbb{F}_{2^2}$, where
$S_0 = 0, S_1 = 1, S_2 = \alpha, S_3 = \alpha+1$. Here, we take 
$P(X) = X^2+X+1$ as the irreducible polynomial to construct
$\mathbb{F}_4$, and $P(\alpha) = 0$ so that $\alpha^2 + \alpha + 1 =
0$.

{\it Initial state:} In Line 1 of algorithm \ref{alg:BFS}, the initial state is
specified by means of a corresponding polynomial $f = \mathcal{F}(S) =
S - 1 - \alpha$. Notice that if we consider the ideal generated by the
initial state polynomial, $I = \langle f\rangle$, then its variety
$V(I) = 1+\alpha$ corresponds to the state encoding $S_3 = \{11\} =
1+\alpha$, where a polynomial in word-level variable $S$ encodes the
initial state. 
% with only one generator $f$, its variety
% $V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 1+\alpha\}$, which equals to $\{1+\alpha\}$, the only
% valid value $S_3$ can take.

@@

{\bf Set operations:} In Lines 5 and 6 of
algorithm  \ref{alg:BFS}, we need  \textbf{union}, 
\textbf{intersection} and \textbf{complement} of varieties over
$\mathbb{F}_{2^k}$, for which we again use algebraic geometry concepts.

@@

In Line 5 of algorithm  \ref{alg:BFS}, we need to compute the complement of a
set of states. Assume that $J$ denotes a polynomial ideal whose
variety $V(J)$ denotes a set of states. We require the computation of
another polynomial ideal $J'$, such that $V(J')=\overline{V(J)}$. We
show that this computation can be performed using the concept of {\bf
  ideal quotient}:

@@

\begin{Proof}
Let $d = gcd(f, g)$. So, $f = df_1 , g = dg_1$ with $gcd(f_1 , g_1 ) =
1$. Note that $f_1 = \frac{f}{gcd(f,g)}$.

@@

Conversely, let $h \in \langle f_1 \rangle$. Then $h = s \cdot f_1$,
where $s \in \Fkk[x]$. So, $hg = hdg_1 = sf_1 dg_1 = sg_1 f \in 
\langle f \rangle$. Therefore, $h \in \langle f : g\rangle$.
\end{Proof}

@@

\begin{Proof}
Since $\Fkk[x]$ is a principal ideal domain, $ J = \langle g\rangle$
for some polynomial $ g \in \Fkk[x]$. Let $d = gcd(g, x^{2^k} -
x)$. So, $g = dg_1 , x^{2^k} - x = df_1$, with $gcd(f_1 , g_1 ) =
1$. Then $J_0 : J = \langle f_1 \rangle$ by Lemma \ref{lem:gcd}.

@@

So, $x \notin V(g)$ and this concludes the proof.
\end{Proof}

@@

With Corollary \ref{cor:complement}, we are ready to demonstrate the
concept of word-level FSM traversal over $\Fkk$ using algebraic
geometry. The algorithm is given in algorithm  \ref{alg:univa}. Note that in
the algorithm, $from^i, to^i, new^i$ are {\it univariate polynomials in variables $S$
or $T$} only, due to the fact that they are the result of a GB
computation with an elimination term order, where the bit-level 
variables are abstracted and quantified away.

@@

\subsection{Word-level FSM Traversal Example}

\begin{Example}
\label{ex:SMPO}
We apply Algorithm \ref{alg:univa} to the example shown in
figure \ref{fig:fsm} to execute the FSM traversal. Let the initial state
$from^0 = \{00\}$ in $\B^2$ or $0 \in \mathbb F_4$. Polynomially, it is
written as $from^0 = S - 0$. In the first iteration, we compose
an ideal $J$ with

@@

\section{Improving our Approach}
\label{sec:improve}
% TODO Revised according to comments in next chapter

Using ATO on computing GB for large set of polynomial is time-consuming, and usually 
intractable. The reason is the exponential computational complexity \cite{gao:gf-gb-ms}:
\begin{Theorem}
Let $J+J_0 = \langle f_1, \dots, f_s, ~x_1^q - x_1, \dots, x_d^q -
x_d\rangle \subset \Fq [x_1, \dots, x_d]$ be an ideal. The time and
space complexity of Buchberger's algorithm to compute a \Grobner
basis of $J+J_0$ is bounded by $q^{O(d)}$.
\end{Theorem}

In our case $q = 2^k$, and when $k$ and $d$ are large, this complexity 
makes abstraction infeasible.

Buchberger's algorithm consists of two major operations: one is finding 
S-poly pairs, the other is dividing the S-poly by the set of polynomials.
Its actual cost is very sensitive to the term ordering: if there exists 
a term order making most S-poly computation unnecessary, then the cost to 
compute S-poly is saved. Moreover, it prevents new polynomials generating 
from unnecessary S-poly reductions, which further reduces the cost
because there are less S-poly pairs as well as candidate divisors.
Besides tuning the term ordering, directly cutting down the number of 
polynomials in the set is also effective.

In order to make our approach scalable, we propose improvements from 
two aspects: on the one hand, using another term ordering which can lower down the 
computational complexity of obtaining GB; on the other hand, reducing the 
polynomials by collecting bit-level primary inputs (PIs) and integrating them as word-level 
variables which are compatible with our working GF.
\subsection{Simplifying the Gr\"obner Basis Computation}
In algorithm \ref{alg:univa}, a Gr\"obner basis is computed for each
iteration to attain the word-level polynomial representation of the next states. In practice, for 
a sequential circuit with complicated structure and large size, Gr\"obner basis computation
is intractable. To overcome the high computational complexity of computing a GB, 
we describe a method that computes a GB of a smaller subset of polynomials.
The approach draws inspirations from \cite{pruss:tcad}. 
According to proposition 2 in \cite{pruss:tcad}, if the GB is computed using {\it refined abstraction term
order} (RATO), there will be only one pair of polynomials $\{f_w,f_g\}$ such that 
their leading monomials are not relatively prime, i.e. 
$$gcd(LM(f_w), LM(f_g)) \neq 1$$
As a well-known fact from Buchberger's algorithm, the S-polynomial 
(Spoly) pairs with relatively prime 
leading monomials will always reduce to 0 modulo the basis and have no contribution to 
the Gr\"obner basis computation.
Therefore, by removing the relatively prime polynomials from $J_{ckt}$, 
the Gr\"obner basis computation is transformed to the reduction of 
$Spoly(f_w,f_g)$ modulo $J_{ckt}$. More specifically, we turn the
GB computation into one-step multivariate polynomial division, and the obtained
remainder $r$ will only contain bit-level inputs and word-level output.

\begin{Example}
After adding intermediate bit-level signal $a,b,c,d$, the elimination ideal for example circuit 
(example \ref{ex:SMPO}) can be rewritten LEX order with RATO:
\begin{align}
&(t_0,t_1)>(a,b,c,d)\nonumber\\&>T>(x,s_0,s_1)\nonumber
\end{align}
We can write down all polynomial generators of $J_{ckt}$:
\begin{equation*}
f_1: a+xs_0s_1+xs_0+xs_1+x+s_0s_1+s_0+s_1+1
\end{equation*}
\vspace{-0.8cm}
\begin{alignat*}{3}
& f_2: b+s_0s_1 &&~~~ f_3: c+x+xs_0 \\
& f_4: d+s_0s_1+s_0 &&~~~ f_5: t_0+ab+a+1 \\
& f_6: t_1+cd+c+d &&~~~ f_7: t_0+t_1\alpha+T 
\end{alignat*}

From observation, the only pair which is not relatively prime is
$(f_5,f_7)$, thus the critical candidate polynomial pair is
$(f_w,f_g)$, where $$f_w = t_0+a\cdot b+a+b, ~~f_g =t_0+t_1\alpha + T$$
Result after reduction is:
\begin{align}
&Spoly(f_w,f_g) \xrightarrow{J+J_0}_{+}T + s_0 s_1 x+\alpha s_0 s_1 \nonumber\\
&+(1+\alpha)s_0 x+(1+\alpha) s_0+s_1 x+s_1+(1+\alpha) x+1\nonumber
\end{align}
The remainder contains only bit-level inputs $(x,s_0,s_1)$ and word-level output $T$.
\end{Example}

The remainder from \emph{Spoly} reduction contains bit-level PS variables, and our objective is to get a polynomial 
containing only word-level PS variables. One possible method is 
to rewrite bit-level variables in term of word-level variables, i.e.
\begin{equation}
\label{eqn:BLVS_reacha}
s_i = \mathcal{G}(S)
\end{equation}

Then we can substitute all bit-level variables with the word-level variable
and obtain a word-level expression.
The authors of \cite{pruss:tcad} propose a method to construct 
a system of equations and solution to the system consists of equation \ref{eqn:BLVS_reacha}.
One way to obtain  is to apply a Gaussian-elimination-fashion approach, which could 
compute corresponding $\mathcal{G}(S)$ efficiently with time complexity $O(k^3)$.

\begin{Example}
{\bf Objective}:\ Abstract polynomial $s_i + \mathcal{G}_i(S)$ from $f_0: s_0+s_1\alpha+S$.

First, compute $f_0^2: s_0+s_1\alpha^2+S^2$. Apparently variable $s_0$ can be
eliminated by operation 
\begin{align}
f_1 =& f_0 + f_0^2 \nonumber\\
=&(\alpha^2+\alpha)s_1+S^2+S\nonumber
\end{align}
Now we can solve univariate polynomial equation $f_1 = 0$ and get solution
$$s_1 = S^2 + S$$
Using this solution we can easily solve equation $f_0 = 0$. The result is
$$s_0 = \alpha S^2+(1+\alpha)S$$
\end{Example}

More formally, polynomial expressions for $s_i$ in terms of $S$ can be
obtained by setting up and solving the following system of equations:
\begin{align}
\label{eqn:betamat}
\begin{bmatrix}
S \\
S^2 \\
S^{2^2} \\
\vdots \\
S^{2^{k-1}}
\end{bmatrix}
&=
\begin{bmatrix}
1 & \alpha & \alpha^{2} & \cdots & \alpha^{{k-1}}\\
1 & \alpha^{2} & \alpha^{4} & \cdots & \alpha^{2(k-1)} \\
1 & \alpha^{4} & \alpha^{8} & \cdots & \alpha^{4(k-1)}\\
\vdots & \vdots & \vdots & \ddots & \vdots \\
1 & \alpha^{2^{k-1}} & \alpha^{2\cdot 2^{k-1}} & \cdots & \alpha^{(k-1)\cdot 2^{k-1}}
\end{bmatrix}
\begin{bmatrix}
s_0\\
s_1\\
s_2\\
\vdots\\
s_{k-1}
\end{bmatrix}
\end{align}
Treat ${\bf s}$ as a vector of $k$ unknowns $s_0,\dots,s_{k-1}$,
then equation \ref{eqn:betamat} can be solved by using Cramer's rule or Gaussian elimination.
In other words, we can obtain $s_i=\mathcal G(S)$ by solving equation \ref{eqn:betamat} symbolically.

In this approach we get word-level variable representation for each bit-level PS variables. 
By substitution, a new polynomial in word-level PS/NS variables could be obtained.

After processing with RATO and bit-to-word conversions, we get a polynomial in the 
form of $f_T = T+\mathcal{F}(S,x)$
denoting the {\bf transition function}. We include a polynomial in $S$ to define the present states $f_S$,
as well as the set of vanishing polynomials for primary inputs 
$J_0^{PI} = \langle x_1^2-x_1,\dots,x_n^2-x_n\rangle$. Using elimination term order with $S>x_i>T$,
we can compute a GB of the elimination ideal $\langle f_T,f_S\rangle + J_0^{PI}$. This GB contains a univariate
polynomial denoting next states. The improved algorithm is depicted in Alg. \ref{alg:refined}.

\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\Indm
 \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\F(S)$}
\Indp

  $from^0 = reached = \F(S)$\;
  $f_T = $Reduce($Spoly(f_w,f_g), J_{ckt})$\;
  \tcc{Compute S-poly for the critical pair, then reduce it with circuit ideal under RATO}
	Eliminate bit-level variables in $f_T$\;
  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
  {
  	$i \gets i + 1$\;
  	$G \gets$GB($\langle f_T , from^{i-1}\rangle+J_0^{PI}$)\;
  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest; $J_0^{PI}$ covers all possible inputs from PIs}
	$to^i \gets G\cup \Fkk[T]$\;
	\tcc{There will be a univariate polynomial in $G$ denoting next state in word-level variable $T$}
	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
	$from^i \gets new^i(S\setminus T)$\;
	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
  }
  \tcc{Loop until a fixpoint reached: newly reached state is empty}
\Return{$\langle reached\rangle$}
\caption {Refined Algebraic Geometry based FSM Traversal}\label{alg:refined}
\end{algorithm}
\DecMargin{1em}

\subsection{PI Partition}
Using above techniques we can get a remainder polynomial with only word-level PS/NS variables. However in most 
cases, the number of bit-level PIs will be too large for the last-step Gr\"obner basis computation. 
Therefore it is necessary to convert bit-level PIs to word-level PI variables.

Assume we have $k$-bit datapath and $n$-bit PIs. In Galois field, we need to carefully partition $n$ PIs
such that states of each partition can be covered by a univariate polynomial respectively.

\begin{Proposition}
Divide $n$-bit PIs to partitions $n_1,n_2,\dots, n_s$ and let $n_1$-bit, $n_2$-bit, $\dots$, $n_s$-bit word-level variables
represent their evaluations in $\mathbb F_{2^k}$ respectively. Then $n_1,n_2,\dots,n_s~|~k$.
\end{Proposition}

Again, assume a partition $n_i~|~k$ and corresponding word-level variable is $P$. Then we can use polynomial
$P^{2^{n_i}}-P$ to represent all signals at free-end PIs, according to following theorem about {\bf composite
field}:
\begin{Theorem}
\label{thm:composite}
Let $k = m\cdot n_i$, such that $\mathbb F_{2^k} = \mathbb F_{(2^{n_i})^m}$. Let $\alpha$ be primitive root of 
$\mathbb F_{2^k}$, $\beta$ be primitive root of the ground field $\mathbb F_{2^{n_i}}$. Then
$$\beta = \alpha^\omega,~~\text{where}~~\omega = \frac{2^k-1}{2^{n_i}-1}$$
\end{Theorem}
\begin{Example}
\label{ex:PI}
In a sequential circuit, PS/NS inputs/outputs are 4-bit signals, which means we will use $\mathbb F_{2^4}$
as working field. PIs are partitioned to 2-bit vectors, which means the ground field is $\mathbb F_{2^2}$.
In ground field we can represent all possible evaluations of this PI partition $\{p_0,p_1\}$ with
$$P^4+P,~~\text{where}~~P=p_0+p_1\cdot\beta$$
Using theorem \ref{thm:composite} we get $\beta = \alpha^5$, so we can redefine word $P$ as element from $\mathbb F_{2^4}$:
$$P = p_0 + p_1\cdot\alpha^5$$
\end{Example}
Using this method we can efficiently partition large size PIs to small number of word-level PI variables.
One limitation of this approach is PIs cannot be partitioned when $k$ is prime.
